Cytokine analysis script - all rounds - adjusted for tissue weight

### 2022-11-26 code/Gene expression.Rmd
updated gene expression

```{r}
delta_gene_expr <- read_excel("data/processed/gene_expression_fold_changes.xlsx") %>%
   mutate(
    groups = recode_factor(  #Change the order of the groups to desired plot order
      groups,
      "con-bf"= "VANCO-BF",
      "sm-form" = "VANCO-SM",
      "fvt-form" = "VANCO-FVF"

      ))
```

@@

## Increased
```{r increased}
increased <- delta_gene_expr %>% 
 group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  filter(mean_fold_change >= 1.75)
```

## Increased plot 
```{r increased plot}
increased_plot <- delta_gene_expr %>% 
  filter(genes %in% increased$genes) %>% 
  mutate(across(genes, factor, levels=c("TNFa","MUC1","IL1b","REG3G"))) %>%
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(relative_fold_change)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_col() +
  facet_wrap(~genes,
             nrow = 1, 
             scales = "free") + 
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.position = "none") + 
    scale_fill_brewer(palette="Set1")
  increased_plot
```

>>>>

```{r}
delta_gene_expr <- read_excel("data/processed/gene_expression_fold_changes.xlsx") %>%
   mutate(
    groups = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "VANCO-BF",
      "SM-FORM" = "VANCO-SM",
      "FVT-FORM" = "VANCO-FVF"
      ))
```

@@

```{r}
delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  ggplot(
    aes(x = groups,
        y = mean_fold_change,
        fill = groups)
  ) + 
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

delta_gene_expr %>% 
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
        y = `2^-(ddct)`,
        fill = groups)
  ) + 
  geom_boxplot() +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )

```{r}
my_comparisons <- list(c("VANCO-FVF", "VANCO-SM"), c("VANCO-BF", "VANCO-FVF"))
```

## Increased
```{r increased}
increased <- delta_gene_expr %>% 
 group_by(genes, groups) %>% 
  summarise(mean_fold_change = mean(`2^-(ddct)`)) %>% 
  filter(mean_fold_change >= 1.75)
```

## Increased plot 
```{r increased plot}
increased_plot <- delta_gene_expr %>% 
  filter(genes %in% increased$genes) %>% 
  mutate(across(genes, factor, levels=c("CXCL1","TNFa","MUC1","IL1b","REG3G"))) %>%
  group_by(genes, groups) %>% 
  ggplot(
    aes(x = groups,
          y = `2^-(ddct)`,
        fill = groups)
  ) +
  #geom_boxplot() +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  facet_wrap(~genes,
             nrow = 1, 
             scales = "free") + 
  labs(
       y = "Fold Δ",
       x = "",
       fill = "Groups") +
  theme_bw()+
  theme(                   
    axis.title.y = element_text(color="darkgrey", size=10, face="bold.italic"),
    axis.ticks.x = element_blank(),  # Hide labels
    axis.text.x = element_blank(),
    legend.position = "none") + 
    scale_fill_brewer(palette="Set1") +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "VANCO-BF",
                   label.y.npc = c(0.55,0.45),
                   label.x = c(3,1))
  #stat_compare_means(comparisons = my_comparisons, label.y.npc = c(0.3,0.5), method = "wilcox.test") 
  increased_plot
```

### 2022-11-26 code/cytokines/Cytokines_tissue_adjusted.Rmd
updated gene expression

#_____Weight of tissue samples_____#
Exp1_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 1) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 1",
    y = "Tissue weight (mg)",
    x = "",
    caption = "Pairwise comparison with CON-BF as reference. \n
    **: p < 0.01, *: p < 0.05, NS: not significant."
  ) 
Exp1_tissue_weight
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp1-weights-of-tissue.png',Exp1_tissue_weight)
```

@@

#_____Weight of tissue samples_____#
Exp2_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 2) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 2",
    y = "Tissue weight (mg)",
    x = "",
    caption = "Pairwise comparison with CON-BF as reference. \n
    **: p < 0.01, *: p < 0.05, NS: not significant."
  ) 
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp2-weights-of-tissue.png',Exp2_tissue_weight)

@@

#_____Weight of tissue samples_____#
Exp3_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 3) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 3",
    y = "Tissue weight (mg)",
    x = "",
    caption = "Pairwise comparison with CON-BF as reference. \n
    **: p < 0.01, *: p < 0.05, NS: not significant."
  ) 
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp3-weights-of-tissue.png',Exp3_tissue_weight)

@@

#_____Weight of tissue samples_____#
Exp4_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 4) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "VANCO-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 4",
    y = "Tissue weight (mg)",
    x = "",
    caption = "Pairwise comparison with VANCO-BF as reference. \n
    ****: p < 0.0005, ***: p < 0.001, NS: not significant."
  ) 
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp4-weights-of-tissue.png',Exp4_tissue_weight)

>>>>

#_____Weight of tissue samples_____#
Exp1_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 1) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 1",
    y = "Tissue weight (mg)",
    x = ""
    #caption = "Pairwise comparison with CON-BF as reference. \n
    #**: p < 0.01, *: p < 0.05, NS: not significant."
  ) 
Exp1_tissue_weight
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp1-weights-of-tissue.png',Exp1_tissue_weight)
```

@@

#_____Weight of tissue samples_____#
Exp2_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 2) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 2",
    y = "Tissue weight (mg)",
    x = ""
    #caption = "Pairwise comparison with CON-BF as reference. \n
    #**: p < 0.01, *: p < 0.05, NS: not significant."
  ) 
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp2-weights-of-tissue.png',Exp2_tissue_weight)

@@

#_____Weight of tissue samples_____#
Exp3_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 3) %>%
  filter(group != "VANCO-BF") %>% 
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "CON-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.5) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 3",
    y = "Tissue weight (mg)",
    x = ""
    #caption = "Pairwise comparison with CON-BF as reference. \n
    #**: p < 0.01, *: p < 0.05, NS: not significant."
  ) 
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp3-weights-of-tissue.png',Exp3_tissue_weight)

@@

#_____Weight of tissue samples_____#
Exp4_tissue_weight <- cytokines_long %>%       
   filter(exp_number == 4) %>%
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  stat_compare_means(aes(label = ..p.signif..),
                   method = "wilcox.test", ref.group = "VANCO-BF",
                   label.y.npc = 0.55) +
  stat_compare_means(method = "kruskal.test",
    aes(label = paste0("Kruskal Wallis Test. \n p = ", ..p.format..)),
    label.y.npc = 0.75,
    label.x.npc = 0.3) +
  geom_point() +
  labs(
    title = "Tissue weight of intestinal samples, experiment 4",
    y = "Tissue weight (mg)",
    x = ""
    #caption = "Pairwise comparison with VANCO-BF as reference. \n
    #****: p < 0.0005, ***: p < 0.001, NS: not significant."
  ) 
ggsave(width = 170, height = 120, units = "mm",'exports/cytokines/Exp4-weights-of-tissue.png',Exp4_tissue_weight)
